Embedded Meta-Learning: Toward more flexible deep-learning models

### 2019-05-03 writing/embedded_meta_learning.tex
Very preliminary poly section

\section{Introduction}
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep[e.g.][]{Lake2016}. One fundamental reason for this is that humans are aware of what we are trying to compute and why. By contrasts, there is a fundamental separation of knowledge within most deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between data and tasks. That is, neural networks knowledge about what is being computed is only implicitly accessible to those computations. \par
There are a number of advantages to representing knowledge about data and tasks. In particular, it can grant the ability to rapidly adapt behavior to a new task. The problem of rapid learning has been partially addressed by meta-learning systems \citep{?????}. However, humans can take our knowledge and flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexiblity. \par
In this paper, we propose a new class of architectures which essentially take a functional perspective on meta-learning and exploit the idea of homoiconicity\footnote{A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.}. By treating tasks and data as functions, we can conceptually think of both data \emph{and} learned behaviors as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to flexibly transform these solutions in response to changing task demands. We suggest that this ability will be key to building more intelligent and flexible deep learning systems. \par
In particular, we propose a meta-learning architecture which embeds data (including natural language inputs) and tasks in a shared space, and meta-learns to execute mappings over this space. We demonstrate that this approach allows flexible remapping of behavior, as well as offering other benefits such as a useful framework for continual learning. \par
\section{Method}
\subsection{Architecture}
The essential idea of the architecture is that we separate the system into two parts: 
\begin{enumerate}
\item Domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$
\item A meta-learning system which:
    \begin{enumerate}
    \item Embeds tasks into the shared embedding space $Z$.
    \item Learns to perform tasks in accordance with their embeddings.
    \end{enumerate}
\end{enumerate}
The utility of having a completely shared space $Z$ in which data, language, and tasks are represented is that it allows for arbitrary mappings between these distinct types of data. In addition to basic tasks, the system could in principle learn to map language to tasks (follow instructions) or tasks to language (explaining behavior), or tasks to tasks (changing behavior). This is a step closer to the flexible reasoning humans display. \par
Without training on these mappings, of course, the system will not be able to execute them well. However, ideally if it is trained on a braod enough set of such mappings, it will be able to generalize these to new instances drawn from the same data/task distribution. For instances that fall outside its data distribution or for optimal performance it may require some retraining. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow. \par

\begin{figure}
\centering
\begin{tikzpicture}[auto]

@@

\section{Experiments}

%\subsection{Boolean functions}
%As a proof of concept, we first evaluated the system on a simple task of computing boolean functions on boolean inputs. Specifically, we can imagine mapping from binary-valued input vectors to a single binary output representing the evaluation of some logical proposition on the input. We train the system to figure out how to compute the predicate from seeing a subset of the (input, output) pairs. \par   
%We can then train the system to do various meta-mappings, for example identifying certain types of functions (like XOR) or negating predicates. \par
%
\subsection{Learning multivariate polynomials}
As a proof of concept, we first evaluated the system on the simple task of learning polynomials in 4 variables of degree $\leq 2$ (i.e. the task was to learn functions of the form $f: \mathbb{R}^4 \rightarrow \mathbb{R}$ where $f \in \mathcal{P}_2 \left(\mathbb{R}\right)$). This yields an infinite family of base-level tasks (the vector space of all such polynomials), as well as many families of meta-mappings over tasks (for example, multiplying a polynomial by a constant, squaring a polynomial, or permuting its input variables). This allows us to not only examine the ability of the system to learn to learn polynomials from data, but also to adapt its learned representations in accordance with these meta-tasks. \par
%% TODO: appendix

@@

In order to switch the system to an RL meta learning system, we made a very simple change. Instead of providing the system with (input, target) tuples to embed, we provided it with (state, action, reward) tuples, and trained it via a DQN-like approach \citep{Mnih2015} to predict rewards for each bet in each state. (A full RL framework is not strictly necessary here because there is no temporal aspect to the tasks; however, we saw it as a useful proof of concept.) The hand is explicitly provided to the network for each example, but which game is being played is implicitly captured in the training examples, without any explicit cues. That is, the system must learn to play directly from seeing a set of (state, action, reward) tuples which implicitly capture the structure and stochasticity of the game. \par   

@@

Our work is an extrapolation from the rapidly-growing literature on meta-learning \citep[e.g.][]{Vinyals2016, Santoro2016, Finn2017a, Finn2018} and continual learning, or more generally tools for avoiding catastrophic interference based on architectural innovations \citep[e.g.][]{Fernando2017, Rusu2016}, regularization terms \citep[e.g.][]{Kirkpatrick2016, Zenke2017} or external memory \citep[e.g.][]{Sprechmann2018}. There has also been interesting recent work on hybrid approaches, such as meta-learning in an online setting \citep{Finn2019}. Our work is specifically inspired by the algorithms that attempt to have the system learn to adapt to a new task via activations rather than gradient descent \citep[e.g.][]{Wang2016a, Duan2016}. \par
Our architecture builds directly off of prior work on HyperNetworks \citep{Ha2016} -- networks which parameterize other networks -- and other recent applications of it, such as guessing parameters for a model to accelerate model search \citep[e.g.][]{Brock2018a}. This is related to the longer history of work on different time-scales of weight adaptation \citep{Hinton1982, Ba2016}. It is more abstractly related to work on learning to propose architectures \citep[e.g.][]{Zoph2016}, and to models that learn to select and compose skills to apply to new tasks \citep[e.g.][]{Andreas, Andreas2016, Tessler2016, Reed2015}. \par 
There has also been some interest in task (or function) embeddings. In simultaneous work \citet{Achille2019} proposed computing embeddings for tasks from the Fisher information of the parameters in a model tuned (final layer only) on the task. They show that this captures some interesting properties of the tasks, including some types of semantic relationships, and can help identify models that can perform well on a task. Other recent work has suggested a similar meta-learning framework where latent codes are computed for a task which can be decoded to a distribution over parameters \citep{Rusu2019}. Our perspective can be seen as a generalization of these that learns task embeddings in an end-to-end way, allows parameterizing a model (while some approaches only allow selection), but most importantly allows for remapping of behavior by meta-tasks. \par

@@

>>>>

\section{Introduction}
Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep[e.g.][]{Lake2016}. One fundamental reason for this is that humans are aware of what we are trying to compute and why. By contrasts, there is a fundamental separation of knowledge within most deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between data and tasks. That is, neural networks knowledge about what is being computed is only implicitly accessible to those computations. \par
There are a number of advantages to representing knowledge about data and tasks. In particular, it can grant the ability to rapidly adapt behavior to a new task. The problem of rapid learning has been partially addressed by meta-learning systems \citep{?????}. However, humans can take our knowledge and flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexiblity. \par
In this paper, we propose a new class of architectures which essentially take a functional perspective on meta-learning and exploit the idea of homoiconicity\footnote{A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.}. By treating tasks and data as functions, we can conceptually think of both data \emph{and} learned behaviors as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to flexibly transform these solutions in response to changing task demands. We suggest that this ability will be key to building more intelligent and flexible deep learning systems. \par
In particular, we propose a meta-learning architecture which embeds data (including natural language inputs) and tasks in a shared space, and meta-learns to execute mappings over this space. We demonstrate that this approach allows flexible remapping of behavior, as well as offering other benefits such as a useful framework for continual learning. \par
\section{Method}
\subsection{Architecture}
The essential idea of the architecture is that we separate the system into two parts: 
\begin{enumerate}
\item Domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system which:
    \begin{enumerate}
    \item Embeds tasks into the shared embedding space $Z$.
    \item Learns to perform tasks in accordance with their embeddings.
    \end{enumerate}
\end{enumerate}
The utility of having a completely shared space $Z$ in which data, language, and tasks are represented is that it allows for arbitrary mappings between these distinct types of data. In addition to basic tasks, the system could in principle learn to map language to tasks (follow instructions) or tasks to language (explaining behavior), or tasks to tasks (changing behavior). This is a step closer to the flexible reasoning humans display. \par
Without training on these mappings, of course, the system will not be able to execute them well. However, ideally if it is trained on a braod enough set of such mappings, it will be able to generalize these to new instances drawn from the same data/task distribution. For instances that fall outside its data distribution or for optimal performance it may require some retraining. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow. \par 
\begin{tikzpicture}[auto]

@@

\section{Experiments}
%\subsection{Boolean functions}
%As a proof of concept, we first evaluated the system on a simple task of computing boolean functions on boolean inputs. Specifically, we can imagine mapping from binary-valued input vectors to a single binary output representing the evaluation of some logical proposition on the input. We train the system to figure out how to compute the predicate from seeing a subset of the (input, output) pairs. \par   
%We can then train the system to do various meta-mappings, for example identifying certain types of functions (like XOR) or negating predicates. \par
%
\subsection{Learning multivariate polynomials}
\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/poly/basic_results.png}
\caption{Basic meta-learning successfully generalizes to held out tasks. The solid line indicates optimal performance.}
\label{poly_basic_results}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/poly/meta_results.png}
\caption{The system generalizes to meta-mapping new tasks. The solid line indicates optimal performance.} 
\label{poly_meta_map_results}
\end{figure}
As a proof of concept, we first evaluated the system on the simple task of learning polynomials in 4 variables of degree $\leq 2$ (i.e. the task was to learn functions of the form $f: \mathbb{R}^4 \rightarrow \mathbb{R}$ where $f \in \mathcal{P}_2 \left(\mathbb{R}\right)$). This yields an infinite family of base-level tasks (the vector space of all such polynomials), as well as many families of meta-mappings over tasks (for example, multiplying a polynomial by a constant, squaring a polynomial, or permuting its input variables). This allows us to not only examine the ability of the system to learn to learn polynomials from data, but also to adapt its learned representations in accordance with these meta-tasks. \par

\textbf{Basic meta learning:} First, we show that the system is able to achieve the basic goal of learning a held-out task from a few examples in Fig. \ref{poly_basic_results}. \par

\textbf{Meta mapping (task $\rightarrow$ task):} Furthermore, the system is able to perform meta-mappings (mappings over tasks) in order to flexibly reconfigure its behavior, as shown in fig. \ref{poly_meta_map_results}. \par

@@

In order to switch the system, we made a very simple change. Instead of providing the system with (input, target) tuples to embed, we provided it with (state, action, reward) tuples, and trained it via a DQN-like approach \citep{Mnih2015} to predict rewards for each bet in each state. (A full RL framework is not strictly necessary here because there is no temporal aspect to the tasks; however, we saw it as a useful proof of concept.) The hand is explicitly provided to the network for each example, but which game is being played is implicitly captured in the training examples, without any explicit cues. That is, the system must learn to play directly from seeing a set of (state, action, reward) tuples which implicitly capture the structure and stochasticity of the game. \par   

@@

Our work is an extrapolation from the rapidly-growing literature on meta-learning \citep[e.g.][]{Vinyals2016, Santoro2016, Finn2017a, Finn2018, Stadie2018} and continual learning, or more generally tools for avoiding catastrophic interference based on architectural innovations \citep[e.g.][]{Fernando2017, Rusu2016}, regularization terms \citep[e.g.][]{Kirkpatrick2016, Zenke2017} or external memory \citep[e.g.][]{Sprechmann2018}. There has also been interesting recent work on hybrid approaches, such as meta-learning in an online setting \citep{Finn2019}. Our work is specifically inspired by the algorithms that attempt to have the system learn to adapt to a new task via activations rather than gradient descent \citep[e.g.][]{Wang2016a, Duan2016}. \par
Our architecture builds directly off of prior work on HyperNetworks \citep{Ha2016} -- networks which parameterize other networks -- and other recent applications of it, such as guessing parameters for a model to accelerate model search \citep[e.g.][]{Brock2018a}. This is related to the longer history of work on different time-scales of weight adaptation \citep{Hinton1982} that has more recently been applied to meta-learning contexts \citep[e.g.][]{Ba2016, Munkhdalai2017}. It is more abstractly related to work on learning to propose architectures \citep[e.g.][]{Zoph2016}, and to models that learn to select and compose skills to apply to new tasks \citep[e.g.][]{Andreas, Andreas2016, Tessler2016, Reed2015}. \par 
There has also been some recent interest in task (or function) embeddings. In simultaneous work \citet{Achille2019} proposed computing embeddings for tasks from the Fisher information of the parameters in a model tuned (final layer only) on the task. They show that this captures some interesting properties of the tasks, including some types of semantic relationships, and can help identify models that can perform well on a task. Other recent work has suggested a similar meta-learning framework where latent codes are computed for a task which can be decoded to a distribution over parameters \citep{Rusu2019}. Our perspective can be seen as a generalization of these that learns task embeddings in an end-to-end way, allows parameterizing a model (while some approaches only allow selection), but most importantly allows for remapping of behavior by meta-tasks. \par

@@

We see our proposal as a logical progression from the fundamental idea of meta-learning -- that there is a continuum between data and tasks. This naturally suggests the idea of manipulating task representations just like we manipulate data.
